---
title: "exploratory_analysis"
author: "mggn"
date: "10/5/2020"
output: html_document
---

## Grouped summaries
quantitative comparisons across groups are informative
group_by() and summarize() are useful in this situation

group_by() adds a layer to your data based on existing variables,
changes behavior of some key functions, not invisible but easy to miss

summarize() allows you to compute one number summaries based on existing variables,
most useful in conjunction with group_by()
produces a dataframe, easy to integrate into a pipeline

```{r setup}
library(tidyverse)

weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```

```{r}
weather_df %>%
  count(month, name = "n_obs")
```

# WARNING: don't use r's table function because it is ugly and unhelpful

```{r eval = FALSE}
weather_df %>%
  pull(month) %>% 
  table
```


```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```
### This is a digression on 2 by 2 tables

case_when allows us to recode variables
(I used this in the last HW lol)

```{r}

weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )
  ) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

```{r}
#reasonable way to create a 2*2 table in R

weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

```{r}

weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))
```
    
```{r}    
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```
  
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```
    
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)

#kable formats results in a way that's easier to read
```

```{r}
  
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point()

```

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax))
```


