---
title: "visualization_lecture"
author: "mggn"
date: "10/5/2020"
output: html_document
---

### Notes pre-coding segment
-exploratory data analysis includes visualization, checks for data completeness
and reliability
-quantification of centrality and variability
-initial evaluation of hypothesis and hypothesis generation
-looking at data is critical
-make dozens of graphics for each dataset

GGPLOT
-makes graphics with relative ease
-cohesiveness, same principles underlie all graphic types
-start google searches with 'ggplot' 
-basic graph components: data, aesthetic mappings, geoms
-advanced graph components: facets, scales, statistics
-combining these components build a graph

### time to code

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```

-rnoaa allows us to pull data from online
  so part one is essentially "bring me that data"
  and then the part in mutate is just data manipulation
  then select to make sure that the order is as I want it
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

### creating scatterplots

```{r}
##new approach same plot
weather_df %>%
  ggplot(aes(x = tmin, y = tmax))+
  geom_point()

```

save and edit a plot object

```{r}
weather_plot =
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

let's get fancy

```{r}

weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)+
  facet_grid(. ~name)
```

## some small notes
How many geoms have to exist?
you can have whatever geoms you want
**but in general it is good to show the data points with geom_point

## univariate plots
Histograms are neat,
we can also make box plots
trendy plots include geom_violin
also ridge plots, aka the most popular plot of 2017 (need to load ggridges package)

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name))+
  geom_histogram()+
  geom_density(alpha = .3)

```

# ggplot part 2

There are some really terrible graphics out there...
to avoid bad plots,
-show as much data as you possibly can
-avoid superfluous thrills
-facilitate comparisons
  put groups in a sensible order, use common axes, use color to highlight 
  groups, NO PIE CHARTS lol
-be clear on intent and your audience
-ggplot is built around a tidy data framework
-touble making a plot could be due to untidy data

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = .5)+
  labs(
    title = "temperature plot",
    x = "min daily temp C",
    y = "max daily temp c",
    caption = "data from rnoaa package; temperatures in 2017"
  )
```

leanring about scales using the same plot

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = .5)+
  labs(
    title = "temperature plot",
    x = "min daily temp C",
    y = "max daily temp c",
    caption = "data from rnoaa package; temperatures in 2017"
  )+
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("15 C", "0", "15 C")
  )+
  scale_y_continuous(
    position = "right"
  )
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name))+
  geom_point(alpha = .5)+
  labs(
    title = "temperature plot",
    x = "min daily temp C",
    y = "max daily temp c",
    caption = "data from rnoaa package; temperatures in 2017"
  )+
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
    )+
  theme(legend.position = "bottom")
  
```

there are a variety of themes that you can use!
like theme_minimal
tweaks to theme need to be added after the theme step